---
title: "Discourse Context Frequency Analysis"
author: "Annie Helms (annie_helms@berkeley.edu)"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    toc: true
    toc_float: true
editor_options: 
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, dev = "cairo_pdf", fig.height = 7, fig.width = 7)
```

Library packages

```{r message = FALSE, warning = FALSE}
library(r2glmm)
library(dplyr)
library(afex)
library(ggplot2)
library(lmerTest) #mixed effects linear regression
library(ggsignif) # helps plot significant relations in ggplot2
library(kableExtra) # formats dataframes in html output
library(emmeans) # tukey post-hoc tests
```

### Upload data set
We will upload a data set containing tokens of /fl-, pl-, kl-/ clusters from La Celestina. Each token is coded for:

- lexical stress
- frequency in text (log frequency in ppm)
- frequency in favorable context (FFC)
- favorability of preceeding phonological context
- favorability of following phonological context
- transmission history (oral, written)
- Modern Spanish realization (preservation, palatalization)
- hapax token (yes if one occurrence, no if multiple)

```{r message = FALSE, echo = FALSE}
setwd("/Users/Annie/Box Sync/PhD/github/dcf_celestina/")
```

```{r}
data = read.csv("data/complete.csv", 
                  stringsAsFactors = TRUE)
head(data, 5) %>%
  kable() %>% # piped to allow for kable formatting
  kable_styling(bootstrap_options = c("striped", # alternates row color in table
                                      "hover", # rows are highlighted when hovered over
                                      "condensed", # row height is condensed
                                      "responsive")) %>% # horizontal scrolling across table
  scroll_box(width = "100%")
```

First we'll group the data so that each word type only contributes one row of data.

```{r}
grouped = data %>%
  distinct(word, .keep_all = TRUE)
```

We'll also filter the data so only word types with more than one instance will be analyzed.

```{r}
grouped_mult = grouped %>%
  filter(hapax == "no")
```

### Logistic Regression Model

We'll run a stepwise fixed effects logistic regression model with the dependent variable of Modern Spanish realization `modern_spa`. The factors that will be included as independent variables in pairwise interactions are: `stress`, `freq_log_ppm`, `hapax`, `FFC`, `transmission`, `fol_fav`, and `latin_stop`.

With a DV with string levels, the reference level will be `preservation`.

```{r results = "hide", warning = FALSE}
test_model = glm(modern_spa ~ stress * freq_log_ppm + stress * FFC + stress * transmission + stress * fol_fav + stress * latin_stop +
                   freq_log_ppm * FFC + freq_log_ppm * transmission + freq_log_ppm * fol_fav + freq_log_ppm * latin_stop +
                   FFC * transmission + FFC * fol_fav + FFC * latin_stop +
                   transmission * fol_fav + transmission * latin_stop +
                   fol_fav * latin_stop, data = grouped_mult, family = "binomial")
step(test_model)
```

The results of stepwise regression indicate that in the model, `fitted probabilities numerically 0 or 1 occurred`. We'll plot transmission history to see how this variable predicts the modern realization.

```{r echo=FALSE}
trans_plot = ggplot(grouped, 
       aes(x = transmission, 
           fill = modern_spa)) + 
  geom_bar(position = "fill",
           width = 0.6) +
  theme_classic() +
  labs(x = "Transmission History",
       y = "Proportion",
       title = "") +
  theme(plot.title = element_text(hjust = 0.5, size = 16),
        axis.title = element_text(size = 12),
        legend.title = element_text(size = 12),
        legend.position = "top") +
  guides(fill = guide_legend(title = "Modern Realization")) +
  scale_fill_hue(labels = c("Palatalization", "Preservation")) +
  scale_x_discrete(labels =c("Oral", "Learned"))
  
trans_plot
```

```{r echo = FALSE, eval = FALSE}
cairo_pdf("transmission.pdf", family = "LM Roman 10", width = 5, height = 4)
trans_plot
dev.off()
```

As the level `learned` categorically occurrs with preservation, we'll get a subset of the data in order to look at what factors predict realization among orally-tranmitted words.

```{r}
zero_one = grouped_mult %>%
  filter(transmission=="oral") %>%
  mutate(dv = case_when(modern_spa =="palatalization" ~ 0,
                        modern_spa =="preservation" ~ 1))
as.integer(zero_one$dv)
```

The factor `fol_fav` is imbalanced across the levels of the DV so we will drop it from the model.

```{r results="hide", warning = FALSE}
test_model_2 = glm(dv ~ freq_log_ppm * FFC * latin_stop, data = zero_one, family = "binomial")
step(test_model_2)
```

```{r}
model_new = glm(dv ~ FFC + latin_stop + FFC:latin_stop, family = "binomial", 
    data = zero_one)
summary(model_new)
```

```{r}
int = exp(0)/(1+exp(0))
prob_highffc = (exp(-4.208e+00)/(1+exp(-4.208e+00)))
prob_lowffc = (exp(-4.208e+00*0.1)/(1+exp(-4.208e+00*0.1)))
int
prob_highffc
prob_lowffc
```

Based on this model, when the FFC is high, the probability of preservation is 1.4%. When the FFC is low, the probability of preservation is 40%. 

```{r echo=FALSE, message=FALSE}
ffc_log_plt = ggplot(zero_one, aes(x=FFC, y=dv)) +
  geom_point() + 
  stat_smooth(method="glm", method.args=list(family="binomial"), se=TRUE) +
  theme_classic() +
  labs(x = "Frequency in Favorable Context (FFC)",
       y = "Probability of Preservation",
       title = "") +
  xlim(0,1) +
  ylim(0,1) +
  theme(plot.title = element_text(hjust = 0.5, size = 16),
        axis.title = element_text(size = 12),
        legend.text = element_text(size = 12),
        legend.title = element_text(size = 14),
        legend.position = c(0.2,0.2),
        legend.background = element_rect(fill = "transparent", colour = NA),
        legend.box.background = element_rect(fill = "transparent", colour = NA))
ffc_log_plt
```

```{r echo = FALSE, eval = FALSE}
cairo_pdf("ffc.pdf", family = "LM Roman 10", width = 5, height = 4)
ffc_log_plt
dev.off()
```

```{r echo = FALSE}
ffc_stop_plot = ggplot(data = zero_one,
                  aes(x = latin_stop,
                      y = FFC,
                      fill = modern_spa)) + 
  geom_boxplot() +
  theme_classic() +
  labs(x = "Latin Cluster",
       y = "FFC",
       title = "") +
  ylim(0,1) +
  guides(fill = guide_legend(title = "")) +
  scale_fill_hue(labels = c("Palatalization", "Preservation"))+
  scale_x_discrete(labels=c("/kl-/", "/fl-/", "/pl-/")) +
  theme(plot.title = element_text(hjust = 0.5, size = 16),
        axis.title = element_text(size = 12),
        axis.text = element_text(size = 10),
        legend.text = element_text(size = 12),
        legend.title = element_text(size = 14),
        legend.position = "top",
        legend.background = element_rect(fill = "transparent", colour = NA),
        legend.box.background = element_rect(fill = "transparent", colour = NA))
ffc_stop_plot
```

```{r echo = FALSE, eval = FALSE}
cairo_pdf("ffc_stop.pdf", family = "LM Roman 10", width = 5, height = 4)
ffc_stop_plot
dev.off()
```

The effect of FFC is strongest in words from the Latin cluster /cl-/.

```{r}
stop_result = grouped_mult %>%
  group_by(latin_stop, modern_spa) %>%
  summarize(count = n())
stop_result
```

